### 2025-10-03 thesis/design/sampling.tex
Split and finish the sampling section

This process enables the embedding space to capture semantic patterns independently
of specific database artifacts and domain terminology, improving the generalizability
across different schemas and domains.

This masking allows \textsc{Natural} to recognize that queries like:

@@

in fact share similar semantic structures, even though they operate on different entities.
Given $\mathcal{W} = \{\text{List}, \text{Show}, \text{all}, \text{the}, \text{with}, \text{name}\}$

the masking would result in:

@@

Embeddings can now be computed for the masked versions of the natural language queries.

@@

Query masking applies similar abstraction principles to the corresponding SQL queries.
This enables \textsc{Natural} to identify SQL patterns and structures.

The SQL masking process transforms all identifiers to \texttt{<mask>} and all
literals to \texttt{<value>}. Thus, the query:

@@

This abstraction enables pattern recognition across different domains while
preserving the essential logical structure of the original SQL statements,
thus focusing on the kind of query, rather than the specific instance.

\subsubsection{Sample Embedding Space – $\mathcal{V}$}

\subsubsection{Schema Distance Index – $\mathcal{D}$}

The schema distance index is a set of id and distance tuples $\{(s_1, d_1), ..., (s_i,d_i)\}$
where $s$ is the database schema name and $d$ is the distance to the
\textit{current} schema. Thus a schema distance index has to be computed for each database schema
\textsc{Natural} is running on.

>>>>

This process enables the embedding space to capture semantic patterns independently
of specific database artifacts and domain terminology, improving the generalizability
across different schemas and domains. Thus given $\mathcal{W} = \{\text{List},
\text{Show}, \text{all}, \text{the}, \text{with}, \text{name}\}$ the natural language
questions:

@@

Will get masked, such that the semantic words are maintained, but domain specific
terminology is discarded:

@@

Now latent space embeddings can be computed for the masked versions of the natural
language questions.

@@

Query masking applies similar abstraction principles to the corresponding SQL queries.
This enables \textsc{Natural} to identify SQL patterns and structures.
The SQL masking process transforms all identifiers to \texttt{<mask>} and all
literals to \texttt{<value>}. Thus, the query:

@@

This abstraction enables pattern recognition across different domains while
preserving the essential logical structure of the original SQL statements.

\subsubsection{Embedding Space}

To construct an embedding space $\mathcal{V}_{\mathcal{S}}$ over a set of samples $\mathcal{S} = \{(q_1, s_1), ..., (q_i, s_i)\}$
where $q$ is the natural language question and $s$ is the corresponding SQL query, we need to compute
the embeddings of $mask(q)$ and $mask(s)$ respectively through an embedding function $\iota$.

Thus the embedding space $\mathcal{V}_{\mathcal{S}}$ can be defined as:

$$\mathcal{V}_{\mathcal{S}} = \{~(q, \iota(mask(q)), s, \iota(mask(s)))~|~(q, s) \in \mathcal{S}~\}$$
